HSC Synthetic Pipeline I: Basic Design and Characterization of the Photometric Performance of HSC

### 2017-04-09 draft/ms_v1.tex
Draft: PSF magnitude accuracies

\def\redm{\texttt{redMaPPer}}
\def\hscpipe{\texttt{hscPipe}}
\def\synpipe{\texttt{SynPipe}}
\def\cmodel{\texttt{cModel}}
\def\forced{\texttt{forced}}
\def\unforced{\texttt{unforced}}
\def\tract{\texttt{Tract}}
\def\visit{\texttt{Visit}}
\def\tracts{\texttt{Tracts}}
\def\visits{\texttt{Visits}}
\def\galsim{\texttt{G}{\scriptsize \texttt{AL}}\texttt{S}{\scriptsize \texttt{IM}}}
\def\hst{{\textit{HST}}}

@@

        \item In Fig \ref{fig:star_sn}, we show the relations between the input
            magnitude and the $\mathrm{S}/\mathrm{N}$ recovered by \hscpipe{}
            photometry in $g$- and $i$-band.

        \item \plan{Define $\mathrm{S}/\mathrm{N}$, quickly explain the purpose.}

        \item \plan{At fixed input magnitude, PSF $\mathrm{S}/\mathrm{N}$ depends 
            on seeing conditions. 
            For $i$-bands, the two \tracts{} have quite different seeing. 
            And bad seeing leads to clearly lower $\mathrm{S}/\mathrm{N}$.}

        \item \plan{Briefly discuss the detection limit of point sources. 
            Basically consistent with the reported values in \citet{HSCDR1}.}

@@

        \item In Fig \ref{fig:psf_mag}, we evaluate the accuracies of \hscpipe{} PSF 
            magnitudes in all five bands.

        \item \plan{The overall performance of PSF photometry is good. 
            In $i$-band, the median magnitude difference is around ${\sim}0.015$ mag
            (${\sim}1.5$\% accuracy) at $i_{\mathrm{Input}}{\sim}19.0$ mag. 
            At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF 
            magnitude is around $3$\%. 
            Down to $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy is 
            around 7\%.
            This is consistent with the external check in \citet{HSCDR1}.}

        \item \plan{At fixed input magnitude, the accuracy does not strongly depends
            on filters. 
            The only exception is the $y$-band, which has slightly worse accuracy. 
            At $y_{\mathrm{Input}}{\sim}23.5$, the average accuracy is already close 
            to $10$\%.}
            \todo{Possible causes? background noise? Worse PSF model?}

       \item \plan{The \forced{} PSF magnitude provides unbiased estimates for
           isolated stars. 
           For $grzy$-bands, the median magnitude difference is ${\sim}0.002$ mag, 
           and can be ignored.}

       \item \plan{Using $i$-band as example, although worse seeing leads to 
           lower $\mathrm{S}/\mathrm{N}$, it does not strongly impact the 
           accuracy of the PSF magnitude. 
           The bad seeing \tract{} suggests almost identical accuracy comparing
           to the good seeing one.}

       \item \plan{However, the story for highly blended stars is quite different. 
           \hscpipe{} systematically under-estimate the total flux of highly blended
           stars by $0.05$--$0.10$ mag at fixed input magnitude.}

       \item \plan{We also test the \unforced{} PSF photometry in all five 
           bands. 
           The results suggest very similar accuracy.
           We highlight the differences between \forced{} and 
           \unforced{} PSF photometry in Fig \ref{fig:psf_diff}. 
           The overall difference is quite small, but it seems that \hscpipe{} 
           \unforced{} PSF photometry recovers slightly more flux at very 
           faint end.} 
           \todo{Move the figure to appendix?}

>>>>

\def\redm{\texttt{redMaPPer}}
\def\hscpipe{\texttt{hscPipe}}
\def\synpipe{\texttt{SynPipe}}
\def\cmodel{\texttt{cModel}}
\def\forced{\texttt{forced}}
\def\unforced{\texttt{unforced}}
\def\tract{\texttt{Tract}}
\def\visit{\texttt{Visit}}
\def\tracts{\texttt{Tracts}}
\def\visits{\texttt{Visits}}
\def\galsim{\texttt{G}{\scriptsize \texttt{AL}}\texttt{S}{\scriptsize \texttt{IM}}}
\def\hst{{\textit{HST}}}
\def\s2n{{$\mathrm{S}/\mathrm{N}$}}

@@

    We first briefly discuss the relationship between input magnitude and the
    signal-to-noise ratio of the output PSF photometry.
    Here, we simply define the \s2n{} as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band
    without considering the aperture correction.
    This will help us visualize
    1) the detection limit for point sources,
    2) the impact of seeing condition on PSF photometric,
    3) and the scatter of photometric accuracies at fixed input magnitude given all
       the realistic features of HSC images (e.g. background subtraction, nearby
       object, optical artefacts).
    We show the results for $g$- and $i$-band in Fig \ref{fig:star_sn}, and find tight
    relations in both bands.

    As expected, we clearly see that the seeing condition impacts the \s2n{} of point
    sources at fixed input magnitude.
    In $i$-band, the \tract{} with slightly worse seeing (0.7\asec{}) shows systematically
    lower \s2n{} comparing to the one with better seeing (0.5\asec{}).
    As for the $g$-band, since the two \tracts{} share similar seeings now, they also
    follow similar relations.
    We also see that the blendedness can also affect accuracies of PSF photometry.
    In $i$-band, the highly blendend synthetic stars scatter toward lower \s2n{} at
    fixed magnitude.
    Interestingly, the effect is not obvious in $g$-band, suggeting the deblending
    process and photometric quality of blended objects also depend on seeing condition.

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$ and
    $i$-band at average seeing condition, which is consistent with the values found
    in \citet{HSCDR1}.
    At the same time, it is worth reminding the users of HSC data that the detection
    limit will show spatial variations due to the seeing conditions.

@@

    Now, we will look into the accuracies of PSF magnitudes in all five bands
    independently.
    In Fig \ref{fig:psf_mag}, we show the relations between the input magnitude and the
    magnitude difference from the \hscpipe{} \forced{} PSF photometry.

    The overall performance of PSF photometry is excellent.
    In $i$-band, the median magnitude difference for \tract{} with 0.5\asec{} seeing is
    around ${\sim}0.015$ mag (${\sim}1.5$\% accuracy) at
    $i_{\mathrm{Input}}{\sim}19.0$ mag.
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the average accuracy of PSF
    magnitud keeps at ${\sim}3$\% level (${\sim}0.036$ mag average difference).
    Down to $i_{\mathrm{Input}}{\sim}25.2$ mag, the average accuracy descreases to
    ${\sim}7$\% level with an average ${\sim}0.072$ mag difference.
    The \tract{} with worse seeing shows similar accuracy of PSF photometry at
    $i_{\mathrm{Input}}<23.5$ mag.
    At faint end, the accuracy becomes slightly worse: ${\sim}5$\% accuracy around
    $i_{\mathrm{Input}}{\sim}24.0$ mag; and ${\sim}11$\% accuracy around
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1 PV2 and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results.
    At fainter magnitude, external comparisons become difficult due to the lack of deep
    imaging data at HSC level and the impact of the color-term when comparing photometry
    using different filters.
    The results from \synpipe{} hence provide more useful evaluations down to the
    detection limit\footnote{Strictly speaking, the uncertainties here should be
    considered as lower limit as the \synpipe{} now does not consider the uncertainties
    in PSF modeling.}.

    We also see that the accuracy of PSF photometry does not strongly depend on filters.
    The forced PSF magnitude in $r$ and $z$-band have very similar accuracy with
    $i$-band (1.5-4.0\% at $19.0 < r_{\mathrm{Input}} < 24.0$ mag, ${\sim}8$\% down to
    $r_{\mathrm{Input}}{\sim}25.2$ mag;
    1.0-5.0\% at $19.0 < z_{\mathrm{Input}} < 24.0$ mag, ${\sim}10$\% down to
    $z_{\mathrm{Input}}{\sim}25.2$ mag).
    The accuracies in $g$- and $y$-band become slightly worse (
    1.5-7.0\% at $19.0 < g_{\mathrm{Input}} < 24.0$ mag, ${\sim}13$\% down to
    $g_{\mathrm{Input}}{\sim}25.2$ mag;
    1.5-5.0\% at $19.0 < y_{\mathrm{Input}} < 23.0$ mag, ${\sim}12$\% down to
    $y_{\mathrm{Input}}{\sim}24.0$ mag;), but the differences can be attributed to the
    differences in seeing conditions.

    For relative isolated stars, \hscpipe{} provides unbiased PSF photometry down to
    very faint magnitude in $i$- and $z$-band.
    For $g$-, $r$-, and $y$-band, we find small offsets in the distributions of magnitude
    differences, which indicates \hscpipe{} tends to underestimate the fluxes of stars
    in these bands by ${\sim}0.01-0.02$ mag.
    Although the exact cause of this offset is unclear (could still be related to
    differenecs in seeing), it will not affect the accuracies of magnitudes of colors.

    At the same time, the story for highly blended stars is very different.
    \hscpipe{} on average systematically underestimates the total fluxes of them by
    $0.05$--$0.10$ mag at fixed input magnitude.
    We see the same effect in all bands, and the impact of blendedness becomes
    increasingly significant at fainter magnitde.
    It is therefore important to keep this in mind when using PSF photometry for point
    sources in HSC data.

    We also test the \unforced{} PSF photometry in all five bands, and find very similar
    accuracies.
    During the measurement of \forced{} PSF photometry, \hscpipe{} fixes the centroid
    of the PSF model across all five band.
    Therefore, the magnitude difference between \forced{} and \unforced{} PSF magnitudes
    indicates the photometric uncertainties induced by differences in the accuracies of
    astrometric calibrations and PSF modeling among different bands.
    We highlight the differences between \forced{} and \unforced{} PSF photometry of
    $g$- and $y$-band in Fig \ref{fig:psf_diff}.
    The overall difference is very small ($<0.01$ mag), but it also suggest that
    \hscpipe{} \unforced{} PSF photometry systematically recovers slightly more fluxes
    at very faint end.
